---
title: "Writing_retreat"
author: "P Wright"
date: "27/01/2022"
output: html_document
---

```{r}
library(plyr)
library(dplyr)
library(rgdal)
library(raster)
library(textshape)

mypath="C:/Users/patrick.wright/Work Folders/Documents/Writing retreat"
```

#56000 lower target Wales
#112000 upper target Wales
#180000 25y target


Data prep

```{r}
score<-raster(paste(mypath, "/england_opportunity_woodland_v4_smart_buffered.tif", sep="")) #dsn="gwc21_overall_score_water_and_nfi_erased", layer="gwc21_overall_score_water_and_nfi_erased")
bng<-crs(score)
#score<-subset(score, grid_code > 23.40 )#only keep scores of 23.5 or more
score[score==0] <- NA


England<-readOGR(dsn=mypath,layer="england")
England<-spTransform(England, bng)#british national grid

# Doesn't need to be repeated aggregating LCM to 100m in Wales
# lcm2020<-raster("C:/Users/patrick.wright/Work Folders/Documents/Writing retreat/lcm-2020-25m_4384650/gb2020lcm25m.tif")
# lcm2020<-crop(lcm2020,extent(Wales))
# lcm2020<-aggregate(lcm2020, 4, fun='max')

extW <- floor(extent(England))#create raster of wales

ScoreRas<-aggregate(score,2, method='max')
plot(ScoreRas)

EnRas <- raster(extW, res=100)
EnRas[is.na(EnRas[])] <- 1
EngRas<-mask(EnRas, England)
crs(EngRas)<-bng#create raster of Wales
Area_England<-cellStats(EngRas, 'sum')#total area of wales in ha

#ScoreRas<-mask(WalRas, score) #create raster of woodland opportunity - Map is already made do not recreate
#ScoreRas<-raster(paste(mypath,"/woodlandOppEngland100m.tif", sep=TRUE)) 
Area_Score<-cellStats(ScoreRas, 'sum')#total area of woodland opportunity in ha

lcm2020<-raster(paste(mypath, "/lcm2020.tif", sep=""))

```


```{r }

#56000 lower target
#112000 upper target
England_target<-230400
input_path <- "C:/Users/patrick.wright/Work Folders/Documents/Writing retreat/Species/"#location of shapefiles
Files <- list.files(input_path, pattern = "[.]shp$", full.names=TRUE)
Files


shps<-lapply(Files, readOGR)


Numb<-c(11:64)#number of shapefiles



Woodland_creationUpper <- lapply(Numb,function(x) {
  #Files1 <- paste(input_path, i, sep="")
  Species <- shps[[x]]
  print("1")
  Species<-spTransform(Species, CRS("+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +units=m +no_defs"))#british national grid
  print("2")
  
  Sp_name<-paste(x)
  print("3")
  
  EngRasSPecies<-mask(EngRas ,Species)#create raster of a sp. distribution
  maxV<-maxValue(EngRasSPecies)#to skip any sp that are not present in the country
  print("Check NAs")
  for(val in maxV){
  if ( is.na(val))    {
    next
  }
  
  print("Check pass")
  
  WoodlandCre_species<-EngRasSPecies%>%
    projectRaster(ScoreRas)%>%
    mask(ScoreRas)#raster of woodland opportunity within sp distribution
print("A")

  crs(WoodlandCre_species)<- bng
  Area_Sp<-cellStats(EngRasSPecies, 'sum')#total area of species in ha
  WoodlandOpp_Sp<-cellStats(WoodlandCre_species, 'sum')#total area of woodland creation for species in ha
  
  
  Areaforplanting<-round((England_target*WoodlandOpp_Sp)/Area_Score)#Estimated amount of woodland planted in the sp distribution

  
  lcmWoodCre_Spe<-lcm2020%>% #land cover map of species od woodland opp inside sp range
    projectRaster(WoodlandCre_species,method='ngb')%>%
    mask(WoodlandCre_species)
  
  lcmWoodCre_Spe[lcmWoodCre_Spe == 1] <- NA #remove any broadleaved woodland

    print("B")

  
  
  x2<-rdply(100,sampleRandom(lcmWoodCre_Spe, Areaforplanting, cells=TRUE) )#select x number of cells (ha) in woodland opportunity area and get land cover value
   print("Progress")
  print("C")

  Hab_count<-x2 %>%
    count(lcm2020)%>%
    mutate(average=n/100)%>%#divided by x number of replicates
    mutate(percentage=(n/100)/Areaforplanting)#divided by x number of replicates
  print("Progress2")
  
  Hab_count<-column_to_rownames(Hab_count, loc = 1)#ugly code to sort results
  Hab_new<-as.data.frame(t(Hab_count))
   
  DF.newN <- Hab_new %>% 
     filter(row.names(Hab_new) %in% c("average"))
  # #colnames(DF.newN)<-Cnames
   row.names(DF.newN) <- NULL
   print("Progress3")
   
  Sum<-cbind(Sp_name ,Area_England, Area_Score,Area_Sp,WoodlandOpp_Sp, Areaforplanting, England_target, DF.newN )
  
  write.csv(Sum, paste(input_path,"Results_Upper_England/", x, "Uppertarget_Res.csv", sep=""))
   print("Bind!")  
  
  #out=bind_rows(out,Sum)
  #print(out)
  
  print("Done!")  

#do.call(bind_rows(Sum))
}})


```




Code to add all results into a single data frame

Summary of column names:
Sp_name - code for species
Species - species name
Area Wales - Number of ha in Wales. Sum of 100m cells in landscape
Area score - Number of ha of Woodland opportunity in Wales
Area_sp - Number of ha of species in country
WoodlandOpp_sp - Number of ha available for species in Area_sp
Areaforplanting - Estimated number of ha planted in species range - assuming planting is evenly distributed in Welsh woodland oportunity area and considering the countries planting target
Wales target - Country target
Other land cover columns - mean number of ha for each variable that will be converted into woodland

```{r}

Res<-list.files(paste(input_path, "Results_Upper_England/", sep=""), pattern = "[.]csv$", full.names=TRUE)
head(Res)
Data<-lapply(Res, read.csv)
Data_com<-do.call("rbind.fill", Data)#as.data.frame(lapply(Data, rbind.fill))
head(Data_com)
#input_path
Data_com<-Data_com%>%
  mutate(Species=case_when(
   Sp_name==1 ~"Apodemus flavicollis",
   Sp_name==2 ~"Apodemus sylvaticus",
   Sp_name==3 ~"Arvicola amphibius",
   Sp_name==4 ~"Barbastella barbastellus",
   Sp_name==5 ~"Capra aegagrus hircus",
   Sp_name==6 ~"Capreolus capreolus",
   Sp_name==7 ~"Castor fiber",
   Sp_name==8 ~"Cervus elaphus",
   Sp_name==9 ~"Cervus nippon",
   Sp_name==10 ~"Crocidura suaveolens",
   
   Sp_name==11 ~"Dama dama",
   Sp_name==12 ~"Eptesicus serotinus",
   Sp_name==13 ~"Erinaceus europaeus",
   Sp_name==14 ~"Felis silvestris",
   Sp_name==15 ~"Glis glis",
   Sp_name==16 ~"Hydropotes inermis",
   Sp_name==17 ~"Lepus europaeus",
   Sp_name==18 ~"Lepus timidus",
   Sp_name==19 ~"Lutra lutra",
   Sp_name==20 ~"Macropus rufogriseus",
   
   Sp_name==21 ~"Martes martes",
   Sp_name==22 ~"Meles meles",
   Sp_name==23 ~"Micromys minutus",
   Sp_name==24 ~"Microtus agrestis",
   Sp_name==25 ~"Microtus arvalis subsp orcadensis",
   Sp_name==26 ~"Muntiacus reevesi",
   Sp_name==27 ~"Mus musculus",
   Sp_name==28 ~"Muscardinus avellanarius",
   Sp_name==29 ~"Mustela erminea",
   Sp_name==30 ~"Mustela nivalis",
   
   Sp_name==31 ~"Mustela putorius",
   Sp_name==32 ~"Myodes glareolus subsp skomerensis",
   Sp_name==33 ~"Myodes glareolus",
   Sp_name==34 ~"Myotis alcathoe",
   Sp_name==35 ~"Myotis bechsteinii",
   Sp_name==36 ~"Myotis brandtii_mystacinus",
   Sp_name==37 ~"Myotis daubentonii",
   Sp_name==38 ~"Myotis nattereri",
   Sp_name==39 ~"Neomys fodiens",
   Sp_name==40 ~"Neovison vison",
   
   Sp_name==41 ~"Nyctalus leisleri",
   Sp_name==42 ~"Nyctalus noctula",
   Sp_name==43 ~"Oryctolagus cuniculus",
   Sp_name==44 ~"Ovis aries",
   Sp_name==45 ~"Pipistrellus nathusii",
   Sp_name==46 ~"Pipistrellus pipistrellus sl",
   Sp_name==47 ~"Pipistrellus pipistrellus sstr",
   Sp_name==48 ~"Pipistrellus pygmaeus",
   Sp_name==49 ~"Plecotus auritus",
   Sp_name==50 ~"Plecotus austriacus",
   
   Sp_name==51 ~"Rattus norvegicus",
   Sp_name==52 ~"Rhinolophus ferrumequinum",
   Sp_name==53 ~"Rhinolophus hipposideros",
   Sp_name==54 ~"Sciurus carolinensis 1995",
   Sp_name==55 ~"Sciurus carolinensis 2005",
   Sp_name==56 ~"Sciurus carolinensis 2010",
   Sp_name==57 ~"Sciurus vulgaris 1995",
   Sp_name==58 ~"Sciurus vulgaris 2005",
   Sp_name==59 ~"Sciurus vulgaris 2010",
   Sp_name==60 ~"Sorex araneus",
   
   Sp_name==61 ~"Sorex minutus",
   Sp_name==62 ~"Sus scrofa",
   Sp_name==63 ~"Talpa europaea",
   Sp_name==64 ~"Vulpes vulpes"
  ))%>%
  rename(#Broadleaved_woodland = X1,
         Coniferous_Woodland = X2,
         Arable_Horticulture = X3,
         Improved_Grassland = X4,
         Neutral_Grassland = X5,
         Calcareous_Grassland = X6,
         Acid_grassland = X7,
         Fen_Marsh_Swamp = X8,
         Heather = X9,
         Heather_grassland = X10,
         Bog = X11,
         Inland_Rock = X12,
         Saltwater = X13,
         Freshwater = X14,
         Supralittoral_Rock = X15,
         Supralittoral_Sediment = X16,
         Littoral_Rock = X17,              ##Warning check if 17 is in your table
         Littoral_Sediment = X18,
         Saltmarsh = X19,
         Urban = X20,
         Suburban = X21
         )  

Data_com[is.na(Data_com)] <- 0

write.csv(Data_com, paste(input_path, "Results_Upper_England/Final_England_Upper.csv", sep=""))


```




Code to add all results into a single data frame

Summary of column names:
Sp_name - code for species
Species - species name
Area Wales - Number of ha in Wales. Sum of 100m cells in landscape
Area score - Number of ha of Woodland opportunity in Wales
Area_sp - Number of ha of species in country
WoodlandOpp_sp - Number of ha available for species in Area_sp
Areaforplanting - Estimated number of ha planted in species range - assuming planting is evenly distributed in Welsh woodland oportunity area and considering the countries planting target
Wales target - Country target
Other land cover columns - mean number of ha for each variable that will be converted into woodland



```{r}

#56000 lower target
#112000 upper target
England_target<-61600
# input_path <- "C:/Users/patrick.wright/Work Folders/Documents/Writing retreat/Species/"#location of shapefiles
# Files <- list.files(input_path, pattern = "[.]shp$", full.names=TRUE)
# Files
# 
# shps<-lapply(Files, readOGR)
# 
# 
 Numb<-c(1:9,11:64)#number of shapefiles


Woodland_creationBiz <- lapply(Numb,function(x) {
  #Files1 <- paste(input_path, i, sep="")
  Species <- shps[[x]]
  print("1")
  Species<-spTransform(Species, CRS("+proj=tmerc +lat_0=49 +lon_0=-2 +k=0.9996012717 +x_0=400000 +y_0=-100000 +ellps=airy +units=m +no_defs"))#british national grid
  print("2")
  
  Sp_name<-paste(x)
  print("3")
  
  EngRasSPecies<-mask(EngRas ,Species)#create raster of a sp. distribution
  maxV<-maxValue(EngRasSPecies)#to skip any sp that are not present in the country
  print("Check NAs")
  for(val in maxV){
  if ( is.na(val))    {
    next
  }
  
  print("Check pass")
  
  WoodlandCre_species<-EngRasSPecies%>%
    projectRaster(ScoreRas)%>%
    mask(ScoreRas)#raster of woodland opportunity within sp distribution
print("A")

  crs(WoodlandCre_species)<- bng
  Area_Sp<-cellStats(EngRasSPecies, 'sum')#total area of species in ha
  WoodlandOpp_Sp<-cellStats(WoodlandCre_species, 'sum')#total area of woodland creation for species in ha
  
  
  Areaforplanting<-round((England_target*WoodlandOpp_Sp)/Area_Score)#Estimated amount of woodland planted in the sp distribution

  
  lcmWoodCre_Spe<-lcm2020%>% #land cover map of species od woodland opp inside sp range
    projectRaster(WoodlandCre_species,method='ngb')%>%
    mask(WoodlandCre_species)
  
  lcmWoodCre_Spe[lcmWoodCre_Spe == 1] <- NA #remove any broadleaved woodland
    print("B")

  
  
  x2<-rdply(100,sampleRandom(lcmWoodCre_Spe, Areaforplanting, cells=TRUE) )#select x number of cells (ha) in woodland opportunity area and get land cover value
   print("Progress")
  print("C")

  Hab_count<-x2 %>%
    count(lcm2020)%>%
    mutate(average=n/100)%>%#divided by x number of replicates
    mutate(percentage=(n/100)/Areaforplanting)#divided by x number of replicates
  print("Progress2")
  
  Hab_count<-column_to_rownames(Hab_count, loc = 1)#ugly code to sort results
  Hab_new<-as.data.frame(t(Hab_count))
   
  DF.newN <- Hab_new %>% 
     filter(row.names(Hab_new) %in% c("average"))
   row.names(DF.newN) <- NULL
   print("Progress3")
   
  Sum<-cbind(Sp_name ,Area_England, Area_Score,Area_Sp,WoodlandOpp_Sp, Areaforplanting, England_target, DF.newN )
  write.csv(Sum, paste(input_path,"Results_BizUS_England/", x, "BizUstarget_Res.csv", sep=""))
   print("Bind!")  
  
  
  print("Done!")  


}})

```



```{r}

Res<-list.files(paste(input_path, "Results_BizUs_England/", sep=""), pattern = "[.]csv$", full.names=TRUE)
head(Res)
Data<-lapply(Res, read.csv)
Data_com<-do.call("rbind.fill", Data)#as.data.frame(lapply(Data, rbind.fill))
head(Data_com)
#input_path
Data_com<-Data_com%>%
  mutate(Species=case_when(
   Sp_name==1 ~"Apodemus flavicollis",
   Sp_name==2 ~"Apodemus sylvaticus",
   Sp_name==3 ~"Arvicola amphibius",
   Sp_name==4 ~"Barbastella barbastellus",
   Sp_name==5 ~"Capra aegagrus hircus",
   Sp_name==6 ~"Capreolus capreolus",
   Sp_name==7 ~"Castor fiber",
   Sp_name==8 ~"Cervus elaphus",
   Sp_name==9 ~"Cervus nippon",
   Sp_name==10 ~"Crocidura suaveolens",
   
   Sp_name==11 ~"Dama dama",
   Sp_name==12 ~"Eptesicus serotinus",
   Sp_name==13 ~"Erinaceus europaeus",
   Sp_name==14 ~"Felis silvestris",
   Sp_name==15 ~"Glis glis",
   Sp_name==16 ~"Hydropotes inermis",
   Sp_name==17 ~"Lepus europaeus",
   Sp_name==18 ~"Lepus timidus",
   Sp_name==19 ~"Lutra lutra",
   Sp_name==20 ~"Macropus rufogriseus",
   
   Sp_name==21 ~"Martes martes",
   Sp_name==22 ~"Meles meles",
   Sp_name==23 ~"Micromys minutus",
   Sp_name==24 ~"Microtus agrestis",
   Sp_name==25 ~"Microtus arvalis subsp orcadensis",
   Sp_name==26 ~"Muntiacus reevesi",
   Sp_name==27 ~"Mus musculus",
   Sp_name==28 ~"Muscardinus avellanarius",
   Sp_name==29 ~"Mustela erminea",
   Sp_name==30 ~"Mustela nivalis",
   
   Sp_name==31 ~"Mustela putorius",
   Sp_name==32 ~"Myodes glareolus subsp skomerensis",
   Sp_name==33 ~"Myodes glareolus",
   Sp_name==34 ~"Myotis alcathoe",
   Sp_name==35 ~"Myotis bechsteinii",
   Sp_name==36 ~"Myotis brandtii_mystacinus",
   Sp_name==37 ~"Myotis daubentonii",
   Sp_name==38 ~"Myotis nattereri",
   Sp_name==39 ~"Neomys fodiens",
   Sp_name==40 ~"Neovison vison",
   
   Sp_name==41 ~"Nyctalus leisleri",
   Sp_name==42 ~"Nyctalus noctula",
   Sp_name==43 ~"Oryctolagus cuniculus",
   Sp_name==44 ~"Ovis aries",
   Sp_name==45 ~"Pipistrellus nathusii",
   Sp_name==46 ~"Pipistrellus pipistrellus sl",
   Sp_name==47 ~"Pipistrellus pipistrellus sstr",
   Sp_name==48 ~"Pipistrellus pygmaeus",
   Sp_name==49 ~"Plecotus auritus",
   Sp_name==50 ~"Plecotus austriacus",
   
   Sp_name==51 ~"Rattus norvegicus",
   Sp_name==52 ~"Rhinolophus ferrumequinum",
   Sp_name==53 ~"Rhinolophus hipposideros",
   Sp_name==54 ~"Sciurus carolinensis 1995",
   Sp_name==55 ~"Sciurus carolinensis 2005",
   Sp_name==56 ~"Sciurus carolinensis 2010",
   Sp_name==57 ~"Sciurus vulgaris 1995",
   Sp_name==58 ~"Sciurus vulgaris 2005",
   Sp_name==59 ~"Sciurus vulgaris 2010",
   Sp_name==60 ~"Sorex araneus",
   
   Sp_name==61 ~"Sorex minutus",
   Sp_name==62 ~"Sus scrofa",
   Sp_name==63 ~"Talpa europaea",
   Sp_name==64 ~"Vulpes vulpes"
  ))%>%
  rename(#Broadleaved_woodland = X1,
         Coniferous_Woodland = X2,
         Arable_Horticulture = X3,
         Improved_Grassland = X4,
         Neutral_Grassland = X5,
         Calcareous_Grassland = X6,
         Acid_grassland = X7,
         Fen_Marsh_Swamp = X8,
         Heather = X9,
         Heather_grassland = X10,
         Bog = X11,
         Inland_Rock = X12,
         Saltwater = X13,
         Freshwater = X14,
         Supralittoral_Rock = X15,
         Supralittoral_Sediment = X16,
         Littoral_Rock = X17,              ##Warning check if 17 is in your table
         Littoral_Sediment = X18,
         Saltmarsh = X19,
         Urban = X20,
         Suburban = X21
         )  

Data_com[is.na(Data_com)] <- 0

write.csv(Data_com, paste(input_path, "Results_BizUs_England/Final_BizUs_England.csv", sep=""))

```